Effect of quality-adjusted production function inputs

### 2015-04-22 Articles/Presentation_ESEE_2015/ESEE 2015.Rmd
Rearranged some of the analysis.

#####################################################
evalCES=FALSE
#####################################################
```

@@

where $y$ is economic output (measured by GDP) and 
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of
capital ($k$), labor ($l$), and energy ($e$).
(The factors of production may be quality-adjusted.)
$time$ is the time variable, usually \texttt{iYear}.

@@

@@

```{r reduce Solow residual PT unadjusted}
ptCDunadjNoE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
                        data = dataPT, constrained = TRUE)
ptCDunadjWithE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
naturalCoef(ptCDunadjWithE)
```

The Solow residual (`lambda`) decreased from 
`r naturalCoef(ptCDunadjNoE)$lambda` to 
`r naturalCoef(ptCDunadjWithE)$lambda`,
so it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the Solow residual significantly.

@@

``` {r reduce Solow residual PT adjusted}
ptCDadjNoE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
                        data = dataPT, constrained = TRUE)
ptCDadjWithE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
naturalCoef(ptCDadjWithE)
```

Using quality-adjusted inputs, the Solow residual (`lambda`) moves from 
`r naturalCoef(ptCDadjNoE)$lambda` to
`r naturalCoef(ptCDadjWithE)$lambda`.

@@

Note that the contribution of capital (`alpha_1`)
always decreases as energy (`alpha_3`) is allowed to play a role.

>>>>

#####################################################
evalCES=TRUE
#####################################################
```

@@

where $y$ is economic output (measured by GDP) and 
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of
capital ($k$), labor ($l$), and energy ($e$).
(The factors of production may be quality-adjusted.)
$time$ is the time variable, usually \texttt{iYear}. 
The \texttt{formula}s are used in a fitting function such as 
\texttt{cdModel} to create a \texttt{model} object.
The function \texttt{naturalCoef} extracts reasonable
coefficients from \texttt{model} objects.

@@

For now, I've organized the analysis around questions that we would like to answer.

@@

```{r}
ptCDunadjNoE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
                        data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
```

```{r}
ptCDunadjWithE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjWithE)
```

The Solow residual (\texttt{lambda}) is always negative. 
However, the absolute value of \texttt{lambda} decreases from 
`r naturalCoef(ptCDunadjNoE)$lambda` to 
`r naturalCoef(ptCDunadjWithE)$lambda`
when adding energy.
So it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the absolute value of the Solow residual somewhat.

@@

``` {r reduce Solow residual PT adjusted}
ptCDadjNoE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
                        data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
```

```{r}
ptCDadjWithE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          data = dataPT, constrained = TRUE)

naturalCoef(ptCDadjWithE)
```

Using quality-adjusted inputs, the Solow residual (`lambda`) moves from 
`r naturalCoef(ptCDadjNoE)$lambda` to
`r naturalCoef(ptCDadjWithE)$lambda` when including energy.
The absolute value of the Solow residual again decreases when adding energy.
But, with quality-adjusted inputs, the effect is much stronger.

@@

Note that the contribution of capital (`alpha_1`)
always shifts to energy (`alpha_3`) when energy is included 
as a factor of production.
